---
title: "Math in Music"
author: "Beckett Hyde & Maedee Trank-Greene"
date: "Due November, 2023"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(tinytex)

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(out.width = "60%")
```



```{r}
# Load in observations_onlydata.csv
observations <- read_csv("observations_onlydata.csv")
summary(observations)
```

```{r}
# scatter plot matrix
colors <- c("#FC4E07", "#00BFFF")

pairs(observations, col = colors[observations$Genre + 1])
```
```{r}
# Want a dataframe with Genre as category, not 0 and 1

observations2 <- observations

observations2$Genre[observations2$Genre == 0] <- "Country"

observations2$Genre[observations2$Genre == 1] <- "Pop"
                                           
ggplot(data = observations2, aes(x = Genre, y = Danceability)) +
  geom_boxplot(fill = 'sky blue') +
  labs(x='Music Genre',y='Danceability') +
  theme_bw() + ggtitle("Boxplot of Music Genre and Danceability")

ggplot(data = observations2, aes(x = Genre, y = Energy)) +
  geom_boxplot(fill = 'sky blue') +
  labs(x='Music Genre',y='Energy') +
  theme_bw() + ggtitle("Boxplot of Music Genre and Energy")

ggplot(data = observations2, aes(x = Genre, y = Tempo)) +
  geom_boxplot(fill = 'sky blue') +
  labs(x='Music Genre',y='Tempo') +
  theme_bw() + ggtitle("Boxplot of Music Genre and Tempo")

ggplot(data = observations2, aes(x = Genre, y = Speechiness)) +
  geom_boxplot(fill = 'sky blue') +
  labs(x='Music Genre',y='Speechiness') +
  theme_bw() + ggtitle("Boxplot of Music Genre and Speechiness")

ggplot(data = observations2, aes(x = Genre, y = Instrumentalness)) +
  geom_boxplot(fill = 'sky blue') +
  labs(x='Music Genre',y='Instrumentalness') +
  theme_bw() + ggtitle("Boxplot of Music Genre and Instrumentalness")

ggplot(data = observations2, aes(x = Genre, y = Liveness)) +
  geom_boxplot(fill = 'sky blue') +
  labs(x='Music Genre',y='Liveness') +
  theme_bw() + ggtitle("Boxplot of Music Genre and Liveness")

ggplot(data = observations2, aes(x = Genre, y = Loudness)) +
  geom_boxplot(fill = 'sky blue') +
  labs(x='Music Genre',y='Loudness') +
  theme_bw() + ggtitle("Boxplot of Music Genre and Loudness")

ggplot(data = observations2, aes(x = Genre, y = Key)) +
  geom_boxplot(fill = 'sky blue') +
  labs(x='Music Genre',y='Key') +
  theme_bw() + ggtitle("Boxplot of Music Genre and Key")

ggplot(data = observations2, aes(x = Genre, y = TimeSignature)) +
  geom_boxplot(fill = 'sky blue') +
  labs(x='Music Genre',y='TimeSignature') +
  theme_bw() + ggtitle("Boxplot of Music Genre and Time Signature")
```


```{r}
# Plot our response variable
plot(observations$Genre,
     main = "Distribution of Response Variable",
     xlab = "Index",
     ylab = "Value",
     yaxt = 'n',
     pch = 16,
     col = colors[observations$Genre + 1])

axis(2, at = c(0, 1))

ratio_pop <- length(observations$Genre[observations$Genre == 1]) / length(observations$Genre)

ratio_country <- length(observations$Genre[observations$Genre == 0]) / length(observations$Genre)
```
From the scatterplot matrix above, we can make some preliminary observations about the relationships between the predictors. 

The following predictors have a weak positive linear relationship: dancebility & energy and energy & tempo. The following predictors have a  

* Positive, linear, weak relationship for dancebility and energy
* Positive, linear, weak relationship for tempo and energy
* 